---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### 1. Load Data
```{r}
act <- read.csv("activity.csv")
names(act)[1] <- "Steps"
names(act)[3] <- "Interval"
```
### 2. Process/Transform the data into a format suitab;e for your analysis
```{r{}}

install.packages("ggplot2")
library(ggplot2)
install.packages("dplyr")
library(dplyr)
install.packages("lattice")
library(lattice)
install.packages("tidyr")
library(tidyr)
install.packages("lubridate")
library(lubridate)
````
### Change Date format to column with date type
```{r}
act$date <- as.Date(act$date , format = "%Y-%m-%d")

act <- act[order(act$date, act$Interval) , ]
names(act)[2] <- "Date"
```
### Create total steps per day
```{r}
act.day <- aggregate(act$Steps, by =list (act$Date), sum)
names(act.day)[1] <- "Date"
names(act.day)[2] <- "Steps"
```
### Create total steps per interval
```{r}
act.int <- aggregate(act$Steps, by=list (act$Interval), sum, na.rm=TRUE, na.action=NULL)
names(act.int)[1] <- "Interval"
names(act.int)[2] <- "Steps"
```
### Create dataframe with with mean steps per interval
```{r}
act.meanint <- aggregate(act$Steps, by=list(act$Interval), mean, na.rm=TRUE, na.action=NULL)
names(act.meanint)[1] <- "Interval"
names(act.meanint)[2] <- "MeanSteps"
```

## What is mean total number of steps taken per day?

### 1. Make a histogram of total number of steps taken each day (ignore missing values)
```{r echo=TRUE}
act.dayNA <- act.day[!is.na(act.day$Steps), ]

png("ACTday.png", width=480, height=480, bg="white")

hist(act.day$Steps, 
  main = "Total Number of Steps Take Each Day",
  xlab= "Total Number of Steps Taken")
  ylab= "Frequency"
  
  dev.off()
```
### 2. Calcutae and report the mean and median totla bumber of steps taken  
  
### Mean number of steps taken each day  
```{r echo=TRUE}
act.mean <- mean(act.day$Steps, na.rm=TRUE)
mean(act.mean)
```

### Median number of steps taken each day
```{r echo=TRUE}
act.median <- median(act.day$Steps, na.rm=TRUE)
median(act.median)
```

## What is the average daily activity pattern?

### 1. Make a Time Series Plot of Average Daily Actvity
```{r echo=TRUE}
png("AVEplot.png", width=480, height=480, bg="white")

plot(act.meanint)

dev.off()
```
### 2. which interval, on average across all the days contains the maximum number of steps

### Which Row and Mean Value
```{r echo=TRUE}
which.max( act.meanint[,2])
max(act.meanint[,2])

```
## Imputing missing values

#### 1. Calculate and report the total number of missing values in the dataset
```{r}
sum(!complete.cases(act))
```
### 2. Fill in the missing values in the dataset
```{r}

act.NA <- merge(act, act.meanint, by = "Interval", sort = FALSE)
## replace NA in Steps with MeanSteps value
names(act.NA)[3] <- "Date"

# Sort data by Date and Interval columns
#act.NA2 <- act.NA[order(act.NA$Date, act.NA$Interval) , ]
act.NA2$Steps <- ifelse(is.na(act.NA2$Steps), act.NA2$MeanSteps, act.NA2$Steps)
act.NA$Steps <- ifelse(is.na(act.NA$Steps), act.NA$MeanSteps, act.NA$Steps)
act.NA$MeanSteps <- NULL
act.NA2$MeanSteps <- NULL
```
### 3. New dataset to reflect no missing values
```{r echo=TRUE}



act.day2 <- aggregate(act.NA$Steps, by =list(act.NA$Date), sum)
names(act.day2)[1] <- "Date"
names(act.day2)[2] <- "Steps"


act.day2 <- aggregate(act.NA2$Steps, by =list(act.NA2$Date), sum)
names(act.day2)[1] <- "Date"
names(act.day2)[2] <- "Steps"

```
### 4. Create a Histogram of Total Number of Steps per day
png("ACTday2.png", width=480, height=480, bg="white")

hist(act.day2$Steps, 
  main = "Total Number of Steps Take Each Day (NA replaced)",
  xlab= "Total Number of Steps Taken")
  ylab= "Frequency"
  
  dev.off()
```
#### 4.B Find Mean and Median total number of steps taken per day
``` {r echo=TRUE}
mean(act.day2$Steps)
median(act.day2$Steps)
```
## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor in th dataset by weekday and weekend
``` {r echo=TRUE}

act.NA2$Weekday2 <-ifelse(weekdays(as.Date(act.day2$Date)) %in% 
                                   c("Saturday", "Sunday"), 
                                "weekend", "weekday")


```

### Make a panel plot containting a time series plot of the 5-minute interval and the average number
### of steps taken averaged across weekday days and weekend days. Use the simulated data.

``` {r echo=TRUE}
act.
act.NA$Weekday <-ifelse(weekdays(act.day2$Date) %in% 
                                   c("Saturday", "Sunday"), 
                                "weekend", "weekday")

act.meanintweekday <- aggregate(act.NA2$Steps, by=list(act.NA2$Interval, act.NA2$Weekday2), mean, na.rm=TRUE, na.action=NULL)
names(act.meanintweekday)[1] <- "Interval"
names(act.meanintweekday)[2] <- "Weekday"
names(act.meanintweekday)[3] <- "MeanSteps"


weekdayplot <- ggplot(act.meanintweekday[act.meanintweekday$weekday == "weekday", ], aes(x = Interval, y = MeanSteps)) + ggtitle("Weekdays")
weekendplot <- ggplot(act.meanintweekday[act.meanintweekday$weekday == "weekend", ], aes(x = Interval, y = MeanSteps)) + ggtitle("Weekends")
```
